Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
pf <- read.csv('/Users/KittusMac/Documents/udacity-eda-course-materials/lesson3/pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response: Majority of the younger age group users have more friend_count compared to older ages.

There is abnormality at 69 and some ages 100+ with higher friend_count. They could be fake accounts/users.

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

ggplot(aes(age, friend_count), data = pf) +
  geom_point() + xlim(13, 90)
  
```

***

### Overplotting
Notes: jitter adds noise and alpha introduces transparency
```{r Overplotting}
ggplot(aes(age, friend_count), data = pf) +
  geom_jitter(alpha = 1/20) + xlim(13, 90)
```

#### What do you notice in the plot?
Response: Young age users have more friends. We can also notice more friend_counts for age 69. Denser points indicate that the majority of the users have lessthan 1000 friend_count

***

### Coord_trans()
Notes: Transforming the y-axis, to get a clear idea on the relationships between the variables.

```{r Coord_trans()}

summary(pf$friend_count)
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(friend_count))) +
  geom_point(alpha = 1/20) +
  coord_trans(y = 'sqrt')

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
summary(pf$friend_count)
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(friend_count))) +
  xlim(13, 90) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  coord_trans(y = 'sqrt')
```

#### What do you notice?

Teenagers have more friend_count. Ages above 90 have more friend_count too
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

ggplot(aes(age, friendships_initiated), 
       data = pf) +
       geom_point(alpha = 1/20, 
          position = position_jitter(height = 0)) +
       xlim(13, 90) + 
       coord_trans(y = 'sqrt')
  
```

Younger ones initiated more requests compared to older ones. Age 69 has initiated more friendship requests strangely
***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library('dplyr')

pf.fc_by_age <- group_by(pf, age) %>% 
                summarise(friend_count_mean = mean(friend_count),
                friend_count_median =  median(friend_count),
                n = n() # counts the number of items and can be used inside summarise functions 
                      ) %>%
                arrange(age)
```

Create your plot!

```{r Conditional Means Plot}

ggplot(aes(x = age, y = friend_count_mean),
       data = pf.fc_by_age) +
       geom_point() +
       scale_x_continuous(breaks = seq(10, 120, 10))

# reveals the pattern by joining the points
ggplot(aes(x = age, y = friend_count_mean),
       data = pf.fc_by_age) +
       geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count),
       data = pf) +
       geom_point(alpha = 1/10, color = 'orange') +
       #coord_trans(y = 'sqrt') +
       geom_line(stat = 'summary', fun.y = mean) +
       geom_line(stat = 'summary', fun.y = median, color = 'blue') +
       geom_line(stat = 'summary', fun.y = quantile, linetype = 2, color = 'blue', fun.args = list(probs = 0.1)) +
       geom_line(stat = 'summary', fun.y = quantile, linetype = 2, color = 'blue', fun.args = list(probs = 0.9)) +
       coord_cartesian(ylim = c(0,1000), xlim = c(0, 70)) 
```

#### What are some of your observations of the plot?
Response: 90% of the users of varies ages between 0 and 90 have friend_count below 1000 except for few ages above 90 who have friend count slightly above 1000

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes: Her histogram plots show the count of people underestimated or overestimated their audience size(in percentages)

***

### Correlation
Notes: Measures the strength between any 2 variables

```{r Correlation}
cor.test(pf$age, pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.027

***

### Correlation on Subsets
Notes: Correlation doesn't imply causality. Descriptive statistics is not sufficient to explain causality. Experimental research (inferential statistics) is needed 

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes: Pearson, Spearman, Kendall

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(y = likes_received, 
           x = www_likes_received), data = pf) +
      geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(y = likes_received, 
           x = www_likes_received), 
      data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'orange')
```
What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response: 0.948

***

### Moira on Correlation
Notes: 

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
data(Mitchell)
str(Mitchell)

ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot. 0.01

b. What is the actual correlation of the two variables?
(Round to the thousandths place)
```{r}
with(Mitchell, cor.test(Month, Temp))
```


```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point() 

```

***

### A New Perspective

What do you notice?
Response: Soil temperatures increased each year and it started decrease after 6th year.

```{r}
?dcor
x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```


Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}


```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))

```

Programming Assignment
```{r Programming Assignment}

pf.fc_by_age_month <- group_by(pf, age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count)) %>%
  arrange(age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_month, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

library(gridExtra)

grid.arrange(p1, p2, ncol = 1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = round(age/5)*5, y = friend_count), 
       data = subset(pf, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean) 
```

***

### Which Plot to Choose?
Notes: All the plots are equally important in exploratory data analysis as each of it may give us different insights. While presenting it to audience we may have to choose the one with import insights.

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

