---
title: 'Problem set: Explore one variables'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exploring Variables in diamonds data set
```{r diamonds}
data(diamonds)
str(diamonds)
```
What letter represents the best color of all the diamonds
```{r}
levels(diamonds$color)
```
Histogram for diamond prices
```{r}
summary(diamonds$price)

qplot(x = price, data = diamonds, binwidth = 100, color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = seq(320, 19000, 500))
```
How many diamonds cost < $500?
```{r}
lessthan500 <- (diamonds$price < 500) 
sum(lessthan500)
```
How many diamonds cost < $250?
```{r}
lessthan250 <- (diamonds$price < 250) 
sum(lessthan250)
```
How many diamonds cost >= $15000?
```{r}
morethaneq15000 <- (diamonds$price >= 15000) 
sum(morethaneq15000)
```
Exploring the largest peak in the histogram
```{r}
qplot(x = price, data = diamonds, xlim = c(620, 630)) 
# there are more diamonds ranging between the price 300 - 1000 
qplot(x = price, data = diamonds) + 
  scale_x_continuous(lim = c(300, 20000), breaks = c(300,20000,1000))

# 675-725 has highest
qplot(x = price, data = diamonds,
      binwidth = 50) +
  scale_x_continuous(lim = c(300, 1000), breaks = seq(300, 1000, 25))

qplot(x = price, data = diamonds,
      binwidth = 1) +
  scale_x_continuous(lim = c(600, 610), breaks = seq(600, 610, 1))
```

histogram of diamond prices by cut
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram() +
  facet_wrap(~cut, ncol = 1)
```

which cut has the highest priced diamond?
```{r}
maxpriceDiamond <- max(diamonds$price)
subset(diamonds$cut, diamonds$price == maxpriceDiamond)
```
which cut has the lowest priced diamond?
```{r}
minpriceDiamond <- min(diamonds$price)
subset(diamonds$cut, diamonds$price == minpriceDiamond)
```
which cut has lowest median price?
```{r}
by(diamonds$price, diamonds$cut, summary)
```
histogram of diamond prices by cut with free scales
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram() +
  facet_wrap(~cut, ncol = 1, scales = 'free')
```
Create a histogram of price per carat and facet it by cut
```{r}
diamonds$pricepercarat <- diamonds$price/diamonds$carat

ggplot(aes(x = pricepercarat), 
       data = diamonds,
       binwidth = 10) +
       geom_histogram() +
       scale_x_log10() +
       facet_wrap(~cut, ncol = 1, scales = 'free')
```
investigate diamond prices by color, clarity and cut
```{r}
#by color
qplot(x = color, y = price, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(300, 8000))

by(diamonds$price, diamonds$color, summary)

#by clarity
qplot(x = clarity, y = price, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(300, 6000))

#median value decreases as the clarity varies except for I1
by(diamonds$price, diamonds$clarity, summary)

#by cut
qplot(x = cut, y = price, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(300, 6500))
by(diamonds$price, diamonds$cut, summary)
```
what is the IQR for the diamonds with best color?
```{r}
by(diamonds$price, diamonds$color, IQR)
```
what is the IQR for the diamonds with worst color?
```{r}
by(diamonds$price, diamonds$color, IQR)
```
investigate diamond pricepercarat by color
```{r}

#best color medians are less than the worst color medians
qplot(x = color, y = pricepercarat, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(300, 6000))

```
investigate the weights of diamonds (carat) using a frequency polygon.
```{r}
qplot(x = carat, data = diamonds, geom = 'freqpoly', binwidth = 0.1, ylim = c(2000, 15000)) +
  scale_x_continuous(limits = c(0, 2), breaks = seq(0.2, 1.5, 0.1))
  #scale_y_continuous(limits = c(2000, 15000)) 
```

